The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-12-15 r_r/letter_to_reviewers.Rmd
updates to manuscript

# Big Changes to the Manuscript.

@@

We agree with all four of these core concerns by Reviewer 2. We sought to develop a revised manuscript that addressed all of these concerns. We concluded that targeted changes to particular parts of the front-end might address some but not all of these concerns. Therefore, we decided to completely destroy the original front end and create a new one. When the dust settled, we were happy to find that the original research questions -- and therefore the original analyses -- remained salient to the new front end.

We explain the structure of the revised manuscript, by drawing from the *Setting the Hook* article by @RN3515, which is my favorite article on the craft of scholarship. They write that "One hook involves using a provocative quotation or vignette to engage the reader in the intriguing and practical nature of their topic" (p. XX). Our revised introduction begins with a vignette describing off-campus recruiting visits to Harpeth Hall, a predominantly white "non-sectarian private school for girls located in the affluent Green Hills neighborhood of Nashville, Tennessee" (p. X). The vignette then moves to the nearby University School of Nashville, which is highly ranked nationally, and receives visits from the more elite private colleges and universities in our sample, but few public universities in our sample. Finally, nearby Father Ryan High School and Christ Presbyterian Academy are lower ranked, are overwhelmingly white and receive visits from few selective private universities in our sample but do receive visits from a substantial share of public research universities in our sample. This introduction relies on data collected for the study, previews some of the core findings, and we hope it "engage(s) the reader in the intriguing and practical nature of the topic" as @RN3515 say.

The revised literature review hopefully sets up a scholarly contribution while also doing a better job of describing the practical motivation for visits to private high schools by public universities. Following @RN3515, our literature review takes the approach of "entering two different conversations and bridging them" with the goal of making a contribution to each conversations while also showing that these two disparate conversations are both parts of a larger story. We first review scholarship on nonresident enrollment by public research universities. Reviewing this scholarship -- which we briefly preview in the introduction -- addresses a core concern from Reviewer 2 by identifying a strong practical motivation for public research universities to target out-of-state private school students (they tend to be affluent and can afford high prices and therefore generate desired tuition revenue). We also show that previous research about recruiting by public research universities has not investigated visits to private high schools.

Second, we review scholarship on linkages between private schools and private universities, we show that prior literature finds a strong relationship between private school attendance and college access and college selectivity. We also present descriptive statistics, which show a strong relationship between private school attendance and attending a selective private college or university. Whereas recent work from economists shows that private school applicants receive higher non-academic ratings than public school applicants [@RN4935], qualitative scholarship by sociologists highlights the importance of social relationships between private schools and selective private universities [@RN4407; @RN3519]. These social relationships are substantially maintained through face-to-face interactions between college admissions officers and high school guidance counselors. These interactions occur during off-campus recruiting visits, motivating the use of social network analyses. Although prior research highlights the importance of social networks between private schools and private universities, they haven't analyzed the social networks that emerge from off-campus visit patterns. Furthermore, we argue that scholarship by sociologists tends to focus on the relationship between elite private schools and elite universities, so we know less about the social networks of private universities that are selective but not elite.

We develop a "bridging" scholarly contribution by stating the these two literatures have been isolated from one another, but they are both parts of a broader story about the changing nature of school segregation and the competition for affluent students. As public universities seek affluent out-of-state students, they are attracted to out-of-state private schools. In doing so, public research universities are encroaching on what had primarily been the territory of private universities. Therefore, analyses of visits to private schools explore the extent to which public and private universities are competing for students at similar or distinct sets of schools, revealing insights about the competition for affluent students. Additionally, prior research shows that private school enrollment is whiter than public schools, that private school students have become more affluent in recent decades, and that private school enrollment has grown in the South as a consequence of "white flight" to private schools. "Thus, in recruiting from private schools, universities may be targeting student bodies that are more affluent and racially homogeneous than public schools. We investigate the characteristics – including geographic region and racial composition – of private schools being targeted by university admissions officers because these characteristics have downstream implications for the racial composition of university campuses."

@@

1. First, we switched the order of RQ2 and RQ3. In the revised manuscript, overlap in recruiting network is RQ2 and characteristics of visited high schools is RQ3. Our rationale is that the characteristics of visited private high schools is the most climactic, substantively important set of results, so we wanted those to go last. Additionally, when thinking through the narrative story of the results, we thought it made more sense to identify overlap -- extent of overlap, which universities visiting similar sets of high schools -- and then examine the characteristics of visited high schools.
1. Second, compared to the original manuscript, the revised manuscript research questions are worded in a way that places less emphasis on comparing public research universities to selective private universities. The original manuscript focused on whether public research universities behave like selective private universities. Here, comparison is central. The revised manuscript is based on the idea both public research universities and selective private universities are targeting the same (or similar) sets of predominantly affluent households. Although comparison remains important, the fundamental interest now becomes developing insights about the competition for students who attend private schools and which schools the universities in our sample are targeting.

@@

>This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affect the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

**Author response**: We wholeheartedly agree with these concern by Reviewer 2 and we are grateful to Reviewer 2 for raising these issues! As discussed above, the revised manuscript addresses these concerns about the practical benefits of recruiting students from private schools. We have also improved the discussion of implications for research and practice.

@@

**Author response**: We agree with his concern. The revised manuscript removes the idea that the mission of public universities is social mobility. As such, we no longer make the strong statement that the historic mission targets low-income state residents as the primary focus.

@@

**Author response**: We wholeheartedly agree with Reviewer 2 here. The previous manuscript's focus on some universally held historic "mission" for public vs. private universities was simultaneously simplistic and distracting. The revised manuscript focuses motivates the analysis based on the more practical idea that universities target private schools because these schools tend to enroll affluent students, who have greater ability to pay high tuition prices without substantial need-based aid.

> **Reviewer concern**. Similarly, I find myself craving some context around the broader implications for this study in the introduction. I think you start to get to this in your conversation around social reproduction, but it would help draw the reader in if you could lay out what the findings of this study could mean for policy and practice.]

@@

> **Reviewer concern**. "The "Chetty" data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile (Burd, 2017; Chetty, Friedman, Saez, Turner, & Yagan, 2020)." P. 3 Instead of saying "Chetty" data, can you be more descriptive here? This sentence assumes a lot of prior knowledge for the reader. "Data linking IRS records to …."

@@

## Section 2 [background and lit review]

@@

**Author response**: The original manuscript had a level-1 section entitled "What do we know about off-campus recruiting?", which reviews market research about off-campus recruiting and published scholarship By contrast, the revised manuscript has a broader literature review section that reviews scholarshp from two literatures -- (1) nonresident enrollment by public universities and (2) linkages between private schools and private universities -- and bridges these two literatures. Therefore, this recommendation from Reviewer 2 is less relevant to the revised literature review structure.

@@

In the revised manuscript, the discussion section substantively describes information about off-campus recruiting events following Covid. In short, off-campus recruiting events are back and are viewed as vital to recruiting efforts. We put this text in the discussion section rather than the literature review because it complements text about recommendations for practice and this text fits less well with the revised literature review structure. Here is the revised text that we placed at the beginning of the "implications for practice" sub-section of the Discussion section.

> Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth ranked expenditure item for public universities and the second ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top ranked item for public universities and the second ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

> The @eab_nacac_counselor_2023 survey of high school counselors asked "how do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business and off-campus recruiting visits remain a vital means of recruiting these students.

@@

**Author response**: We appreciate this concern and agree with the recommendation. We thought of several tables and figures that could help readers visualize some of the network concepts. However, we made the decision to exclude most of these because the manuscript is already quite long. The revised manuscript does include a visualization of the full two-mode recruiting network between universities and high schools (CRYSTAL STATE FIGURE NUMBER ONCE YOU ADD IT). The manuscript also includes a visualization of the ego network of University of Notre Dame (Figure 1), which is useful in explaining what are ego networks and their relationship with RQ3 (characteristics of high schools that reveive visits)

@@

The reason the neighboring state analysis doesn't quite work is exemplified by the University of Nebraska Lincoln (UNL). Nebraska does not share a border with Illinois but a huge proportion of their high school recruiting visits -- public and private -- are in Chicago. The same is true for the University of Arkanas and visits to Georgia (Atlanta), CU-Boulder and California and CU-Boulder and Chicago, and UC-Berkeley and visits to Washington (Seattle). To a huge degree, recruiting visits occur in major metropolitan areas, so the analysis you recommend bears fruit for universities in states that neighbor a state with a major metopolitan area (e.g., University of Alabama and Georgia, or University of Pittsburgh and New York).

We considered an alternative to your analysis that focuses on visits to metropolitan areas within some distance (e.g., 500 miles) from the University. But this analysis is also not clean because distances seem to be greater in the West (e.g., Denver to Los Angeles) than the east coast (e.g., Amherst MA to NYC). For these reasons, we opted to exclude those analyses even though we do agree with the basic logic the Reviewer lays out.

@@

**Author response**: The revised Data and Sample section states that "The rationale for different criteria for inclusion across different university types is because these criteria reflected the interests of a foundation that funded the data collection." We agree with Reviewer 1 that making comparisons between public and private universities would be more appropriate if we used the same classification method to define the data collection sample for all universities. The revised manuscript now includes a Limitations subsection that describes this limitation and its consequences.

@@

**Author response**: This is an excellent suggestion by Reviewer 1. The new Limitations section in the revised manuscript discusses the drawbacks of a convenience sample.

@@

**Author response**: We added text to the Data and Sample section to address this concern. For public research universities, we double checked with the admissions offices and issued formal public records requests if the university did not provide data.

@@

**Author response**: The revised manuscript now states that we used Python to scrape admissions websites [CRYSTAL -- ADD WHICH PACKAGES WE USED]. Additionally, in case this is helpful to readers, the revised manuscript states that all analyses were conducted using *R* and network analyses were conducted using the *igraph* package. As an aside, we could have done the web-scraping in *R* too.

@@

The results are well-organized and clear

@@

**Author response**: We agree with Reviewer 2 implied concern here that just because the privatization research did not address this topic, it is not clear that conducting this analysis makes a significant scholarly contribution because the privatization literature is less relevant than it was 10 years ago.

@@

In the "implications for scholarship" sub-section of the revised Discussion section, we discuss the contribution of this manuscript to the two respective literatures. Next, we argue that our broader/more important contribution is showing that these literatures are part of the same story and developing new insights about the competition privileged students. Here, I paste text from the revised manuscript:

@@

> Our analyses shed light about the nature of this competition. When making decisions about which private schools to visit, public research universities seem to be mimicking the behavior of particular private universities rather than mimicking the behavior of private universities writ large. Several public universities seem to be competing directly for students targeted by a particular private university. For example, UC-Berkeley visited 27.9% of the schools visited by Northwestern, 28.9% of the schools visited by Tufts, and 27.8% of the schools visited by Emory (Figure \@ref(tab:privu-percent-visited-by-pubu-matrix)). We reason UC-Berkeley is competing directly for students because all four universities are in the same broad echelon of prestige. Other potential examples of direct competition include the University of Georgia targeting schools visited by Emory, SMU, TCU, and Tulane, or the University of Arkansas targeting schools visited by Baylor.

> However, public universities that made the greatest number of private school visits seemed to be targeting lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievement, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

**Author response**: As stated previously, the revised manuscript discusses implications for student enrollment composition and university administrators (which we might think of senior administrators, high-level enrollment management administrators). We do not state implications that are explicitly for lower-level admissions professionals. Page limits were a consideration here. Additionally, based on our prior conversations with enrollment managers and admissions officers we believe that junior admissions professionals have limited discretion over which schools a university visits. However, if the Reviewer is adamant that the manuscript should be further revised to address implications for a particular group, we will gladly take this on in a subsequent revision.

@@

Additionally, we follow Reviewer 1's recommendation of removing the level-1 section "Social Network Analysis Concepts" text, and allocating this text to relevant sub-sections of the revised "Methods" section. This change reduces redundancy in text, helps to reduce word count, and results in a structure that we hope is easier for readers to digest. The revised organization of the level-1 Methods section is as follows: Data and sample; Constructing social networks; Analyses (which are described separately by research question)

@@

**Author response**: The reviewer is making a good point here. The one-mode network of university recruiting networks with universities colored by cluster indicates which universities are similar or disparate from one another in terms of private schools visited, as do the tabular counts of overlap (Tables 5 and 6). However, these results don't actually show how the recruiting strategies differ across/within public and private universities in terms of what sorts of private schools they visit. This limitation is one reason that we chose to switch the order of RQ2 and RQ3. In the revised manuscript, we show the analyses of network similarity and then we examine -- separately for each university -- the characteristics of private schools that receive visits. We hope this change helps address the concern. We considered subsequent analyses that would examine the characteristics of private schools that received visits from one set of universities but not another set of universities (e.g., schools visited by Midwestern private universities but not Midwestern puboic universities). There are many potential analyses that could be conducted here, and the choice of which analyses to include felt somewhat arbitrary. Therefore, we decided to exclude such analyses. Since comparison between public and private universities is less of a focus than the initial manuscript, we hope this omisssion is ok.

@@

However, we did include a new analysis that examines the racial composition of schools in the South that received visits. We did this for both visits to private schools and visits to public schools. Our rationale for this comparison is that Universities in our sample tended to visit private schools in the South more than private schools in other regions (often this was true even for universities not located in the South). Additionally, over the last two decades, private school enrollment growth has been strongest in the South and scholars have argued that "white flight" from desegregated public schools is one reason for this enrollment growth [@RN4504]. Therefore, we were interested in examining the racial composition of visited Southern private schools and how this compares to the racial composition of visited Southern public schools.

@@

**Author response**: As stated above, we agree with the concern that our analysis sample is a convenience sample. The revised limitations section states this explicitly and discusses why this limitation undermines the external validity of our findings. We agree with Reviewer 1 that because we do not have longitudinal data, we cannot make strong claims about which university is mimicking which other university. We added this limitation to the Limitations section

@@

> Although a handful of public research universities have relied on nonresident enrollment for many decades (e.g., University of Michigan, University of Delaware), Public research universities began substantially increasing nonresident enrollment en masse around 2005 [@RN3753]. By contrast, selective private universities have long relied on enrollment from private schools [@RN3519; @RN4640]. Therefore, the pursuit of out-of-state private school students by public research universities is relatively recent  encroachment on the historic territory of selective private universities and in-state flagships.

>>>>

# Big Changes to the Manuscript

@@

We agree with all four of these core concerns by Reviewer 2. We sought to develop a revised manuscript that addressed all of these concerns. We concluded that targeted changes to particular parts of the front end might address some but not all of these concerns. Therefore, we decided to completely destroy the original front end and create a new one. When the dust settled, we were happy to find that the original research questions -- and therefore the original analyses -- remained salient to the new front end.

We explain the structure of the revised manuscript, by drawing from the *Setting the Hook* article by @RN3515, which is my favorite article on the craft of scholarship. They write that "One hook involves using a provocative quotation or vignette to engage the reader in the intriguing and practical nature of their topic" (p. 873)<!--https://journals.aom.org/doi/epdf/10.5465/amj.2011.4000-->. Our revised introduction begins with a vignette describing off-campus recruiting visits to Harpeth Hall, a predominantly white "nonsectarian private school for girls located in the affluent Green Hills neighborhood of Nashville, Tennessee" (p. 1). The vignette then moves to the nearby University School of Nashville, which is highly ranked nationally, and receives visits from the more elite private colleges and universities in our sample, but few public universities in our sample. Finally, nearby Father Ryan High School and Christ Presbyterian Academy are lower ranked, overwhelmingly white, and receive visits from few selective private universities in our sample but do receive visits from a substantial share of public research universities in our sample. This introduction relies on data collected for the study, previews some of the core findings, and we hope it "engage(s) the reader in the intriguing and practical nature of the topic" as @RN3515 say.

The revised literature review hopefully sets up a scholarly contribution while also doing a better job of describing the practical motivation for visits to private high schools by public universities. Following @RN3515, our literature review takes the approach of "entering two different conversations and bridging them" (p. 874) with the goal of making a contribution to each conversation while also showing that these two disparate conversations are both parts of a larger story. We first review scholarship on nonresident enrollment by public research universities. Reviewing this scholarship -- which we briefly preview in the introduction -- addresses a core concern from Reviewer 2 by identifying a strong practical motivation for public research universities to target out-of-state private school students (they tend to be affluent and can afford high prices and therefore generate desired tuition revenue). We also show that previous research about recruiting by public research universities has not investigated visits to private high schools.

Second, we review scholarship on linkages between private schools and private universities. We show that prior literature finds a strong relationship between private school attendance and college access and college selectivity. We also present descriptive statistics, which show a strong relationship between private school attendance and attending a selective private college or university. Whereas recent work from economists shows that private school applicants receive higher non-academic ratings than public school applicants [@RN4935], qualitative scholarship by sociologists highlights the importance of social relationships between private schools and selective private universities [@RN4407; @RN3519]. These social relationships are substantially maintained through face-to-face interactions between college admissions officers and high school guidance counselors. These interactions occur during off-campus recruiting visits, motivating the use of social network analyses. Although prior research highlights the importance of social networks between private schools and private universities, they haven't analyzed the social networks that emerge from off-campus visit patterns. Furthermore, we argue that scholarship by sociologists tends to focus on the relationship between elite private schools and elite universities, so we know less about the social networks of private universities that are selective but not elite.

We develop a "bridging" scholarly contribution by stating that these two literatures have been isolated from one another, but they are both parts of a broader story about the changing nature of school segregation and the competition for affluent students. As public universities seek affluent out-of-state students, they are attracted to out-of-state private schools. In doing so, public research universities are encroaching on what had primarily been the territory of private universities. Therefore, analyses of visits to private schools explore the extent to which public and private universities are competing for students at similar or distinct sets of schools, revealing insights about the competition for affluent students. Additionally, prior research shows that private school enrollment is whiter than public schools, that private school students have become more affluent in recent decades, and that private school enrollment has grown in the South as a consequence of "white flight" to private schools. "Thus, in recruiting from private schools, universities may be targeting student bodies that are more affluent and racially homogeneous than public schools. We investigate the characteristics -- including geographic region and racial composition -- of private schools being targeted by university admissions officers because these characteristics have downstream implications for the racial composition of university campuses" (p. 12).

@@

1. First, we switched the order of RQ2 and RQ3. In the revised manuscript, overlap in recruiting network is RQ2 and characteristics of visited high schools is RQ3. Our rationale is that the characteristics of visited private high schools is the most climactic, substantively important set of results, so we wanted those to go last. Additionally, when thinking through the narrative story of the results, we thought it made more sense to identify overlap -- extent of overlap, which universities visiting similar sets of high schools -- and then examine the characteristics of visited high schools.
1. Second, compared to the original manuscript, the revised manuscript research questions are worded in a way that places less emphasis on comparing public research universities to selective private universities. The original manuscript focused on whether public research universities behave like selective private universities. There, comparison is central. The revised manuscript is based on the idea that both public research universities and selective private universities are targeting the same (or similar) sets of predominantly affluent households. Although comparison remains important, the fundamental interest now becomes developing insights about the competition for students who attend private schools and which schools the universities in our sample are targeting.

@@

>This manuscript yields practical implications for public and private universities concerned about racial and socioeconomic diversity. Following the Covid-19 Pandemic, off-campus recruiting visits have bounced back because they remain a fundamental means of maintaining relationships with high school guidance counselors [@RN4896; @RN4895; @eab_nacac_counselor_2023]. Prior research finds that students are sensitive to recruiting interventions, including off-campus recruiting visits [@RN4324; @RN4402]. Therefore, which high schools a university chooses to visit affects the subsequent enrollment composition of the university. In the aftermath of the Supreme Court decision on race-conscious admissions, public and private universities are searching for ways to increase racial diversity without running afoul of the law. The @RN4954[p. 9] recommends that universities expand "outreach to high schools with substantial populations of low-income students and students of color." Most public and private universities in our sample visited a disproportionate number of private schools and these private schools tended to enroll mostly white students. We recommend that universities concerned about racial diversity should conduct an internal equity audit of their recruiting practices, for example, comparing the racial and socioeconomic characteristics of schools they visit versus schools they do not visit.

@@

**Author response**: We wholeheartedly agree with these concerns by Reviewer 2 and we are grateful to Reviewer 2 for raising these issues! As discussed above, the revised manuscript addresses these concerns about the practical benefits of recruiting students from private schools. We have also improved the discussion of implications for research and practice.

@@

**Author response**: We agree with this concern. The revised manuscript removes the idea that the mission of public universities is social mobility. As such, we no longer make the strong statement that the historic mission targets low-income state residents as the primary focus.

@@

**Author response**: We wholeheartedly agree with Reviewer 2 here. The previous manuscript's focus on some universally held historic "mission" for public vs. private universities was simultaneously simplistic and distracting. The revised manuscript motivates the analysis based on the more practical idea that universities target private schools because these schools tend to enroll affluent students, who have greater ability to pay high tuition prices without substantial need-based aid.

> **Reviewer concern**. Similarly, I find myself craving some context around the broader implications for this study in the introduction. I think you start to get to this in your conversation around social reproduction, but it would help draw the reader in if you could lay out what the findings of this study could mean for policy and practice.

@@

> **Reviewer concern**. "The 'Chetty' data on parental income by college show that the overwhelming majority of students at selective private institutions have parental income in the top quintile (Burd, 2017; Chetty, Friedman, Saez, Turner, & Yagan, 2020)." P. 3 Instead of saying "Chetty" data, can you be more descriptive here? This sentence assumes a lot of prior knowledge for the reader. "Data linking IRS records to …."

@@

## Section 2 (background and lit review)

@@

**Author response**: The original manuscript had a level-1 section entitled "What do we know about off-campus recruiting?", which reviews market research about off-campus recruiting and published scholarship. By contrast, the revised manuscript has a broader literature review section that reviews scholarship from two literatures -- (1) nonresident enrollment by public universities and (2) linkages between private schools and private universities -- and bridges these two literatures. Therefore, this recommendation from Reviewer 2 is less relevant to the revised literature review structure.

@@

In the revised manuscript, the discussion section substantively describes information about off-campus recruiting events following Covid. In short, off-campus recruiting events are back and are viewed as vital to recruiting efforts. We put this text in the discussion section rather than the literature review because it complements text about recommendations for practice and this text fits less well with the revised literature review structure. Here is the revised text that we placed at the beginning of the "Implications for Practice" sub-section of the Discussion section:

> Market research suggests that off-campus recruiting visits will remain a vital means of connecting to students and guidance counselors, just as they were prior to the Covid pandemic. @RN4664 found that "travel" was the top-ranked marketing/recruiting expenditure item for both public universities and private universities. During the height of Covid, off-campus recruiting visits ceased and were partially replaced by virtual counselor events [@parents_covid_rnl]. Admissions counselors are now traveling again. @RN4896 found that travel was the fifth-ranked expenditure item for public universities and the second-ranked expenditure item for private universities. When asked about areas of future expenditure growth, travel was the top-ranked item for public universities and the second-ranked item for private universities [@RN4896], suggesting that universities are increasing travel following the decline of the pandemic.

> The @eab_nacac_counselor_2023 survey of high school counselors asked, "How do you prefer to receive information from colleges?" High school visits by college representatives was by far the highest rated "channel" for private school counselors and was also the highest rated channel for public school counselors. By contrast, virtual visits by admissions representatives received low ratings by high school guidance counselors. Finally, @RN4895 asked universities to rank the most effective "outreach strategies for high school counselors." "Calling after visits" was deemed effective for 97% of private universities (the top strategy) and was deemed effective by 88% of public universities. What insights do we glean from these market research findings? In the era of mass marketing and data science, the pursuit of traditional college students from privileged schools and communities is still a hand-to-hand business, and off-campus recruiting visits remain a vital means of recruiting these students.

@@

**Author response**: We appreciate this concern and agree with the recommendation. We thought of several tables and figures that could help readers visualize some of the network concepts. However, we made the decision to exclude most of these because the manuscript is already quite long. The revised manuscript does include a visualization of the full two-mode recruiting network between universities and high schools (Appendix Figure 1). The manuscript also includes a visualization of the ego network of University of Notre Dame (Figure 1), which is useful in explaining what are ego networks and their relationship with RQ3 (characteristics of high schools that receive visits).

@@

The reason the neighboring state analysis doesn't quite work is exemplified by the University of Nebraska-Lincoln (UNL). Nebraska does not share a border with Illinois but a huge proportion of their high school recruiting visits -- public and private -- are in Chicago. The same is true for the University of Arkanas and visits to Georgia (Atlanta), CU-Boulder and California and CU-Boulder and Chicago, and UC-Berkeley and visits to Washington (Seattle). To a huge degree, recruiting visits occur in major metropolitan areas, so the analysis you recommend bears fruit for universities in states that neighbor a state with a major metropolitan area (e.g., University of Alabama and Georgia, or University of Pittsburgh and New York).

We considered an alternative to your analysis that focuses on visits to metropolitan areas within some distance (e.g., 500 miles) from the University. But this analysis is also not clean because distances seem to be greater in the West (e.g., Denver to Los Angeles) than the East Coast (e.g., Amherst MA to NYC). For these reasons, we opted to exclude those analyses even though we do agree with the basic logic the Reviewer lays out.

@@

**Author response**: The revised Data and Sample section states that "The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection." We agree with Reviewer 1<!--TODO: Is this comment by Reviewer 1 or 2? This is under the level-1 section "Response to Reviewer 2"--> that making comparisons between public and private universities would be more appropriate if we used the same classification method to define the data collection sample for all universities. The revised manuscript now includes a Limitations subsection that describes this limitation and its consequences.

@@

**Author response**: This is an excellent suggestion by Reviewer 1<!--TODO: Is this comment by Reviewer 1 or 2? This is under the level-1 section "Response to Reviewer 2"-->. The new Limitations section in the revised manuscript discusses the drawbacks of a convenience sample.

@@

**Author response**: We added text to the Data and Sample section to address this concern. For public research universities, we double-checked with the admissions offices and issued formal public records requests if the university did not provide data.

@@

**Author response**: The revised manuscript now states that we used Python to scrape admissions websites using the *requests* library. Additionally, in case this is helpful to readers, the revised manuscript states that all analyses were conducted using *R* and network analyses were conducted using the *igraph* package. As an aside, we could have done the web-scraping in *R* too.

@@

> **Reviewer comment**. The results are well-organized and clear

@@

**Author response**: We agree with Reviewer 2's implied concern here that just because the privatization research did not address this topic, it is not clear that conducting this analysis makes a significant scholarly contribution because the privatization literature is less relevant than it was 10 years ago.

@@

In the "Implications for Scholarship" sub-section of the revised Discussion section, we discuss the contribution of this manuscript to the two respective literatures. Next, we argue that our broader/more important contribution is showing that these literatures are part of the same story and developing new insights about the competition for privileged students. Here, I paste text from the revised manuscript:

@@

> Our analyses shed light on the nature of this competition. When making decisions about which private schools to visit, public research universities seem to be mimicking the behavior of particular private universities rather than mimicking the behavior of private universities writ large. Several public universities seem to be competing directly for students targeted by a particular private university. For example, UC-Berkeley visited 27.9% of the schools visited by Northwestern, 28.9% of the schools visited by Tufts, and 27.8% of the schools visited by Emory (Table 5). We reason UC-Berkeley is competing directly for students because all four universities are in the same broad echelon of prestige. Other potential examples of direct competition include the University of Georgia targeting schools visited by Emory, SMU, TCU, and Tulane, or the University of Arkansas targeting schools visited by Baylor.

> However, public universities that made the greatest number of private school visits seemed to be targeting the lower-achieving students at private schools targeted by particular private universities. For example, the University of Alabama visited 52.2% of the 625 private schools that Notre Dame visited, with most of the overlap at Catholic schools. Given the differences in academic prestige between them, the University of Alabama may not be competing for the same students as Notre Dame. Instead, they seem to be targeting Catholic school students of a similar economic echelon, who have lower academic achievements, and want to attend a football powerhouse. Additionally, CU-Boulder visited 42.4% of schools visited by Northwestern, which could be interpreted as CU-Boulder competing for lower-achieving students of a similar economic and cultural echelon at the schools targeted by Northwestern. Similarly, the University of South Carolina competes heavily at schools targeted by Emory, SMU, TCU, and Tulane.

@@

**Author response**: As stated previously, the revised manuscript discusses implications for student enrollment composition and university administrators (which we might think of as senior administrators, high-level enrollment management administrators). We do not state implications that are explicitly for lower-level admissions professionals. Page limits were a consideration here. Additionally, based on our prior conversations with enrollment managers and admissions officers, we believe that junior admissions professionals have limited discretion over which schools a university visits. However, if the Reviewer is adamant that the manuscript should be further revised to address implications for a particular group, we will gladly take this on in a subsequent revision.

@@

Additionally, we follow Reviewer 1's recommendation of removing the level-1 section "Social Network Analysis Concepts" text, and allocating this text to relevant sub-sections of the revised "Methods" section. This change reduces redundancy in text, helps to reduce word count, and results in a structure that we hope is easier for readers to digest. The revised organization of the level-1 Methods section is as follows: Data and Sample; Constructing Social Networks; Analyses (which are described separately by research question).

@@

**Author response**: The reviewer is making a good point here. The one-mode network of university recruiting networks with universities colored by cluster indicates which universities are similar or disparate from one another in terms of private schools visited, as do the tabular counts of overlap (Tables 5 and 6). However, these results don't actually show how the recruiting strategies differ across/within public and private universities in terms of what sorts of private schools they visit. This limitation is one reason why we chose to switch the order of RQ2 and RQ3. In the revised manuscript, we show the analyses of network similarity and then we examine -- separately for each university -- the characteristics of private schools that receive visits. We hope this change helps address the concern. We considered subsequent analyses that would examine the characteristics of private schools that received visits from one set of universities but not another set of universities (e.g., schools visited by Midwestern private universities but not Midwestern pubic universities). There are many potential analyses that could be conducted here, and the choice of which analyses to include felt somewhat arbitrary. Therefore, we decided to exclude such analyses. Since comparison between public and private universities is less of a focus than the initial manuscript, we hope this omission is okay.

@@

However, we did include a new analysis that examines the racial composition of schools in the South that received visits. We did this for both visits to private schools and visits to public schools. Our rationale for this comparison is that universities in our sample tended to visit private schools in the South more than private schools in other regions (often this was true even for universities not located in the South). Additionally, over the last two decades, private school enrollment growth has been strongest in the South and scholars have argued that "white flight" from desegregated public schools is one reason for this enrollment growth [@RN4504]. Therefore, we were interested in examining the racial composition of visited Southern private schools and how this compares to the racial composition of visited Southern public schools.

@@

**Author response**: As stated above, we agree with the concern that our analysis sample is a convenience sample. The revised Limitations section states this explicitly and discusses why this limitation undermines the external validity of our findings. We agree with Reviewer 1 that because we do not have longitudinal data, we cannot make strong claims about which university is mimicking which other university. We added this limitation to the Limitations section.

@@

> Although a handful of public research universities have relied on nonresident enrollment for many decades (e.g., University of Michigan, University of Delaware), public research universities began substantially increasing nonresident enrollment en masse around 2005 [@RN3753]. By contrast, selective private universities have long relied on enrollment from private schools [@RN3519; @RN4640]. Therefore, the pursuit of out-of-state private school students by public research universities is a relatively recent encroachment on the historic territory of selective private universities and in-state flagships.

# References
